Cultural Transmission and the Social Contexts of Technological Transitions during the Late Paleolithic of Korea

### 2023-04-26 analysis/paper/paper.qmd
Edited conclusion

In comparing assemblages with more than five stemmed points, we found complex variation in CV values that may suggest mixtures of biases in the transmission of tool-making techniques. We did not find strong support for distinctive, exclusive cultural transmission biases in tool-making in different assemblages. Correlation coefficient values across the assemblages confirm the overall chronological trend indicating indirect bias in Phase 3. Raw material choice appears to be important in driving this trend, with acid volcanic rocks and porphyry becoming more abundant in Phase 3, and dominating assemblages at Yonsangdong that date to this period. One possibility is that these more easily manageable raw materials were part of an adaptive shift during the Last Glacial Maximum, where perhaps the cooler climate favored an expanded range for searching for raw materials, leading to a higher diversity. Similarly, changes in occupation and mobility patterns may have resulted in more frequent contacts and connections with members of social networks [@park2022did]. This may have provided more opportunities to learn point-making directly from socially successful individuals, and increased the contribution of indirect bias during cultural transmission during Phase 3.

>>>>

In comparing assemblages with more than five stemmed points, we found complex variation in CV values that may suggest mixtures of biases in the transmission of tool-making techniques. We did not find strong support for distinctive, exclusive cultural transmission biases in tool-making in different assemblages. Correlation coefficient values across the assemblages confirm the overall chronological trend indicating indirect bias in Phase 3. Raw material choice appears to be important in driving this trend, with acid volcanic rocks and porphyry becoming more abundant in Phase 3, and dominating assemblages at Yonsangdong that date to this period. One possibility is that these more easily manageable raw materials were part of an adaptive shift during the Last Glacial Maximum, where perhaps the cooler climate favored an expanded range for searching for raw materials, leading to a higher diversity. In our previous research, we observed that forager groups with stemmed points were able to occupy higher elevation, where the temperature was generally lower [@park2022did]. The changes in occupation and mobility patterns may have resulted in more frequent contacts and connections with members of increased social networks. This may have provided more opportunities to learn point-making directly from socially successful individuals, and increased the contribution of indirect bias during cultural transmission during Phase 3.
